Reproducible Research - Peer Assessment 1
========================================================

We first set the working dicrectory to be `setwd("C:/Users/pc2/SkyDrive/Coursera/Data Science/Reproducible Research")`.

We read in the data and then determine the total number of steps taken each day.

```{r}
activity <- read.csv("activity.csv",header=TRUE,sep=",")
noofstepsbyday <- sapply(split(activity$steps,activity$date),sum)
```

We next plot a histogram for the above data.

```{r fig.width=7, fig.height=6}
hist(as.numeric(noofstepsbyday),main="Total number of steps taken each day",xlab="Total number of steps")
```
```{r echo=FALSE}
meannoofsteps <- round(mean(noofstepsbyday,na.rm=TRUE),1)
mediannoofsteps <- median(noofstepsbyday,na.rm=TRUE)
```

The mean of the number of steps taken each day is `r meannoofsteps` and the median of the steps taken each day is `r mediannoofsteps`.

```{r}
activity$intervalsteps <- as.factor(activity$interval)
noofstepsbyinterval <- aggregate(steps ~ intervalsteps, activity, mean)
```

A plot showing the average number of steps taken at each interval of the day is shown below.

```{r fig.width=7, fig.height=6}
plot(noofstepsbyinterval$intervalsteps,noofstepsbyinterval$steps,type="l",main="Avg number of steps during a day",xlab="Time of day",ylab="Average number of steps")
```

```{r echo=FALSE}
maxavgsteps <- max(noofstepsbyinterval$steps)
timeofdaywhenmaxstepstaken <- noofstepsbyinterval[noofstepsbyinterval$steps==maxavgsteps,][1,1]
```

The maximun number of steps taken in a 5 minute interval begins at `r timeofdaywhenmaxstepstaken` in the morning.

Imputing missing values
----------------------------------
```{r}
setwd("C:/Users/pc2/SkyDrive/Coursera/Data Science/Reproducible Research")
activity <- read.csv("activity.csv",header=TRUE,sep=",")
cc <- activity[complete.cases(activity),]
TotalMissingValues <- nrow(activity) - nrow(cc)
```

The number of rows in the original data set is `r nrow(activity)` and of these `r TotalMissingValues` have missing values.

We replace the missing value for the specific time interval with the mean for that time interval as calculated above. We then create a new data set that contains the replaced values instead of missing values.

```{r}
notcompletecases <- activity[!complete.cases(activity),]
temp <- merge(notcompletecases,noofstepsbyinterval,by.x="interval",by.y="intervalsteps")
notcompletecases$steps <- temp$steps.y
completecases <- activity[complete.cases(activity),]
activity_new <- rbind(completecases,notcompletecases)
```

For this dataset with the imputed values, we plot a histogram as shown below.

```{r fig.width=7, fig.height=6}
noofstepsbyday_new <- sapply(split(activity_new$steps,activity_new$date),sum)
hist(as.numeric(noofstepsbyday_new),main="Total number of steps taken each day",xlab="Total number of steps")
```
```{r echo=FALSE}
meannoofsteps_new <- round(mean(noofstepsbyday_new,na.rm=TRUE),1)
mediannoofsteps_new <- median(noofstepsbyday_new,na.rm=TRUE)
```
After replacing the missing values with the mean of the values for those specific intervals, the mean of the number of steps taken each day is `r meannoofsteps_new` and the median of the steps taken each day is `r mediannoofsteps_new`.

We want to find differences in activity patterns between weekdays and weekends. To develop this analysis, we add a new column to decide if a particular date is a weekday or weekend. Then we create a new data set that calculates the mean number of steps taken in each 5 minute interval of the day.

```{r}
activity_new$day <- weekdays(as.Date(activity_new[,"date"]),abbreviate=TRUE)
weekends <- subset(activity_new,day=="Sat"|day=="Sun")
weekdays <- subset(activity_new,day=="Mon"|day=="Tue"|day=="Wed"|day=="Thu"|day=="Fri")
weekends$weekend <- "weekend"
weekdays$weekend <- "weekday"
activity_new <- rbind(weekdays,weekends)
meanactivitysteps <- aggregate(steps ~ interval + weekend,data=activity_new,mean)
```
The following panel plot shows the distribution of the average number of steps across time on weekdays and weekends.
```{r fig.width=7, fig.height=6}
library(lattice)
xyplot(meanactivitysteps$steps ~ meanactivitysteps$interval | meanactivitysteps$weekend,layout=c(1,2),type="l",xlab="Time of day",ylab="Mean number of steps")
```


